### 2013-04-29 article.tex
Proof read manuscript
Signed-off-by: Michael Barton <mail@michaelbarton.me.uk>

The increasing accessibility and reduced costs of sequencing has made genome
analysis accessible to more and more researchers. Yet there remains a steep
learning curve in the subsequent computational steps required to process raw
sequence reads into a database-deposited genome sequence. Here we describe
``Genomer,'' a tool to simplify the manual tasks of finishing and uploading a
genome sequence to a database. Genomer can format a genome scaffold into the
common files required for submission to GenBank. This software also simplifies
updating a genome scaffold by allowing a human-readable YAML format file to be
edited instead of large sequence files. Genomer is written as a command line
tool and is an effort to make the manual, and increasingly common, process of
genome scaffolding more robust and reproducible. Extensive documentation and
video tutorials are available at http://next.gs.

@@

Manually editing a genome sequence or annotation is a non-trivial problem
requiring effort on a researcher's part. The large size of FASTA sequence files
or general feature format (GFF) annotation files makes changing these in a text
editor difficult. Tasks such as organising project files, adding additional
sequences to an assembly, updating incorrect annotations or preparing files for
database deposition all divert effort from the main goal of analysing a genome.
As a result there is a steep learning curve for researchers embarking on genome
sequencing for the first time, while still requiring effort from experienced
researchers.

@@

Genomer is written as a command-line tool for use in shells such as bash or
zsh. The advantage of command-line tools over GUI-based tools, is that the former
are simpler to automate and allow the sharing of steps with other researchers
working on related projects. For instance, during our microbial genome project
we automated Genomer using GNU Make build files, allowing the finishing steps
to be repeated automatically using the \verb+make+ command. Genomer may
therefore be of particular interest to bioinformaticians coordinating in groups
on the same genome project and those who prefer automated approaches and
scripting on the command line.

@@

Genomer was written to satisfy our needs for building and finishing megabase
sized draft genomes where assembly of the draft sequence takes place in memory.
This makes Genomer well-suited for microbial genome projects where
approximately 99\% of microbial genomes sequenced to date are less than 10MBp.
The Assembly of very large sequences may however limit performance when the
size of the sequence exceeds available. We however anticipate that genome sizes
of even 3GBp will fit in the memory of a modern laptop or desktop computer.

One features of Genomer is the simplification of generating the annotation
table file required for submission to GenBank. Genomer requires only a GFF3
file containing gene-type entries be required. Genomer is able to generate the
additional encoded protein features from these entries. One caveat to this
however is that the assumption there is no intron/exon structure and therefore
users wishing to submit annotations with alternative splicing cannot take
advantage of this Genomer feature.

>>>>

The increasing accessibility and reduced costs of sequencing has made genome
analysis accessible to more and more researchers. Yet there remains a steep
learning curve in the subsequent computational steps required to process raw
sequence reads into a database-deposited genome sequence. Here we describe
``Genomer,'' a tool to simplify the manual tasks of finishing and uploading a
genome sequence to a database. Genomer can format a genome scaffold into the
common files required for submission to GenBank. This software also simplifies
updating a genome scaffold by allowing a human-readable YAML format file to be
edited instead of large sequence files. Genomer is written as a command line
tool and is an effort to make the manual process of genome scaffolding more
robust and reproducible. Extensive documentation and video tutorials are
available at http://next.gs.

@@

Manually editing a genome sequence or annotation is non-trivial and requires
effort on a researcher's part. The large size of FASTA sequence files or
general feature format (GFF) annotation files makes changing these in a text
editor difficult. Tasks such as organising project files, adding additional
sequences to an assembly, updating incorrect annotations or preparing files for
database deposition all divert effort from the main goal of analysing a genome.
As a result there is a steep learning curve for researchers embarking on genome
sequencing for the first time, while still requiring effort from experienced
researchers.

@@

Genomer is written as a command-line tool for use in shells such as bash or
zsh. The advantage of command-line tools over GUI-based tools, is that the
former are simpler to automate and allow the sharing of scripts with other
researchers working on related projects. For instance, during our microbial
genome project we automated Genomer using GNU Make build files, allowing the
finishing steps to be repeated automatically using the \verb+make+ command.
Genomer may therefore be of particular interest to bioinformaticians
coordinating in groups on the same genome project and those who prefer
automated approaches and scripting on the command line.

@@

Genomer was written to satisfy our needs for building and finishing megabase
sized draft genomes where assembly of the sequence takes place in memory. This
makes Genomer well-suited for microbial genome projects where approximately
99\% of microbial genomes sequenced to date are less than 10MBp. The assembly
of very large sequences may however limit performance when the size of the
sequence exceeds available RAM space. We however anticipate that genome sizes
of even 3GBp will fit in the memory of a modern laptop or desktop computer.

One features of Genomer is the simplification of generating the annotation
table file required for submission to GenBank. Genomer requires only a GFF3
file containing gene-type entries. Genomer is then able to generate the
additional encoded protein features from these entries. One caveat to this
however is that the assumption there is no intron/exon structure and therefore
users wishing to submit annotations with alternative splicing cannot take
advantage of this Genomer feature.
